\documentclass[11pt]{scrartcl}
\usepackage{graphicx} % Required for inserting images
\usepackage[sexy,hints]{evan}
\usepackage{asymptote}

\title{OG! Harmonic Motion Basics.}
\author{OG Sanskar OG Gupta}
\date{March 2025}

\begin{document}

\maketitle

\section{Introduction}
OG!
\section{Problems}
\begin{problem*}
     At what maximum value of $N$ can each unit square of an $N × N$ chessboard be painted red or blue in such a way that the four unit squares at the intersection of any two columns and any two rows of this board are not all the same colour?
\end{problem*}
  
\paragraph{Solution:}
  We claim the answer is $n=4$, achieved by the construction

\begin{asy}
import graph;
size(200);

string rows[4] = {"BBRR", "BRBR", "RBRB", "RRBB"};

for (int i = 0; i < 4; ++i) {
  for (int j = 0; j < 4; ++j) {
    char cell = rows[i][j];
    pen cellColor = (cell == 'B' ? blue : red);
    filldraw(
      (j, 3 - i) -- 
      (j+1, 3 - i) -- 
      (j+1, 4 - i) -- 
      (j, 4 - i) -- 
      cycle,
      cellColor,
      black
    );
  }
}

for (int i = 0; i <= 4; ++i) {
  draw((0,i) -- (4,i));
  draw((i,0) -- (i,4));
}
\end{asy}
\textbf{Proof that this is the maximum:}
Assume for $N \geq 5$. We consider the bottom most row, By PHP, there are at least the 3 squares of the same colour. WLOG, let it be blue. Let these 3 squares collectively be called $OG-cells$.\\\\ Now we completely focus on the 3 columns that contain the three squares , since $N \geq 5$, there are atleast $5$ rows. Label the rows from bottom to top as $1 ... n$. Now, Row one contains 3 blue squares, exactly above these 3 blue squares(meaning in the same column as the 3 squares) in any row(except row 1)there must be at maximum 1 blue square( $as$ $if$ $there$ $are$ $2$ $blue$ $squares$ $,$ $then$ the intersections of that row and row 1 with respective columns of the 2 blue squares would all be blue). Hence in any other row there are 2 red 1 blue or all 3 red squares exactly above the $OG-cells$ in the bottom row. 
Now since $N \geq 5$, therefore we have at least 4 other rows other than row 1 i.e. Row 2,3,4,5 exist.\\\\

\textbf{\textit{Case 1:}} If in any of Row $2,3,4,5$ the three squares above the $OG-cells$ are all red\\\\
WLOG, let the row be ROW 2, then In $Row 3$, two of the three squares above the $OG-cells$, must be of the same colour(By PHP). \\\\However this is not possible, as then we would group $2 -red$ squares with any 2 of the cells of ROW 2 which are above the $OG-cells$ and $2-blue$ squares with any 2 of the $OG-cells$. 

\textbf{\textit{Case 2:}} In none of Row $2,3,4,5$ the three squares above the $OG-cells$ are all red,\\\\
then in each of $ROW 2,3,4,5$, in the three squares above the $OG-cells$, 2 of them are $red$ and 1 is $blues$. \\We know that there are 3 permutations of $2 Reds$ and $1 Blue$, but we have $4$ rows($2,3,4$ and $5$). \\
Hence in at least 2 rows there must be the same permutation of the 3 squares above the $OG-cells$, but then we would have 2 rows and 2 columns whose intersection squares all are red.
    
